---
title: "Economic_Sentiment_Analysis"
author: "Diveesha Vanchireddy"
format:
  html:
    date: today
    toc: true
    code-overflow: wrap
    code-fold: true 
    embed-resources: true
    df-print: paged
---


## Data Setup: 
1) How to pull bds2022_st_cty (now bds2023_st_cty) from pubdata? - got it. 
2) How to merge BDS and CHRR using FIPS as composite key? - paste st X county but CHHR also has a FIPS code already so just going ot use that. 
3) How to align years between BDS and CHRR? average CHHR 2022 = 2019-2020 surveyed data. 
4) What about 2019, 2020, 2021 CHRR files for panel?

## EDA tasks listed: 
1) Descriptive stats of all variables (mean, median, sd, min, max)
2) Correlation matrix of all variables to see initial relationships
3) Time series plots of BDS reallocation rate and CHRR health outcomes over time
4) Scatter plots of reallocation rate vs mental health and physical health outcomes
5) Statistical tests comparing coefficients of economic distress in mental vs physical health models
6) Analysis of BDS pandemic-era volatility (2019–2022) to see if there were significant changes during COVID years
7) Robustness checks using other BDS measures of job volatility (job creation rate, job destruction rate, etc.) to see if results hold across different specifications


# Main analytical questions to answer: 
1) Mental unhealthy days vs physically unhealthy days — WHICH VARIABLE TO FOCUS ON?
2)Is mental health more associated with economic distress than physical health? Do a stat test comparing coefficient magnitudes
3) Correlation between insured vs uninsured adults in terms of health impacts?
4) Does reallocation rate work as the main explanatory variable?
5) Reallocation rate = default explanatory variable



# Next steps:
1) Aligning CHRR data.
```{r}
library(pubdata)
library(tidyverse)
get_merged_cross_section <- function(bds_year, chrr_offset = 2) {
  chrr_year <- as.character(bds_year + chrr_offset)
  chrr_raw <- pubdata::get("chrr", chrr_year)
  my_vars <- c(
    "v001_rawvalue", "v036_rawvalue", "v042_rawvalue", "v145_rawvalue",
    "v024_rawvalue", "v023_rawvalue", "v082_rawvalue", "v085_rawvalue",
    "v009_rawvalue", "v011_rawvalue"
  )
  
  chrr_wi <- chrr_raw |>
    filter(statecode == "55", countycode != "000", variable %in% my_vars) |>
    select(fipscode, county, variable, value) |>
    pivot_wider(names_from = variable, values_from = value) |>
    rename(
      premature_death = v001_rawvalue,
      phys_unhealthy = v036_rawvalue,
      mental_unhealthy = v042_rawvalue,
      freq_mental_distress = v145_rawvalue,
      child_poverty = v024_rawvalue,
      unemployed = v023_rawvalue,
      single_parent = v082_rawvalue,
      uninsured = v085_rawvalue,
      smokers = v009_rawvalue,
      obese = v011_rawvalue
    )
  
 bds_slice <- bds_wi |>
    filter(year == bds_year) |>
    select(fipscode, reallocation_rate, job_creation_rate, job_destruction_rate,
           net_job_creation_rate, estabs_entry_rate, estabs_exit_rate)
 result <- inner_join(chrr_wi, bds_slice, by = "fipscode") |>
    mutate(bds_year = bds_year, chrr_release = as.integer(chrr_year))
  return(result)

}

```
```{r}
bds_raw <- pubdata::get("bds", "2022_st_cty") # swtich to 2023_st_cty becuase more recent and pandemic era data. 
bds_wi <- bds_raw |>
  filter(st == "55") |>
  mutate(
    fipscode = paste0(st, cty),
    year = as.integer(year),
    across(
      c(job_creation_rate, job_destruction_rate, net_job_creation_rate,
        reallocation_rate, estabs_entry_rate, estabs_exit_rate),
      ~as.numeric(replace(.x, .x == "X", NA))
    )
  )
```
```{r}
# Pre-COVID cross-section
pre_covid <- get_merged_cross_section(2019)

# Post-COVID cross-section
post_covid <- get_merged_cross_section(2022)

#get both
both <- bind_rows(pre_covid, post_covid)
both |>
  select(fipscode, bds_year, premature_death:obese) |>
  pivot_longer(premature_death:obese) |>
  ggplot(aes(x = value, fill = factor(bds_year))) +
  geom_histogram(bins = 20, alpha = 0.6, position = "identity") +
  facet_wrap(~name, scales = "free") +
  labs(fill = "BDS Year")
  

# Check what you got
cat("Pre-COVID:", nrow(pre_covid), "counties\n")
cat("Post-COVID:", nrow(post_covid), "counties\n")
```



## Performing EDA on the merged dataset:
- We have successfully merged the BDS and CHRR datasets using FIPS codes, creating a comprehensive dataset for analysis.
## getting descriptive stats for all variables in this one cross-section dataset:

```{r}
desc_stats <- both |>
  select(bds_year, premature_death:obese, reallocation_rate,
         job_creation_rate, job_destruction_rate, net_job_creation_rate) |>
  pivot_longer(-bds_year, names_to = "variable", values_to = "value") |>
  group_by(bds_year, variable) |>
  summarise(
    n = sum(!is.na(value)),
    mean = round(mean(value, na.rm = TRUE), 3),
    sd = round(sd(value, na.rm = TRUE), 3),
    min = round(min(value, na.rm = TRUE), 3),
    p5 = round(quantile(value, 0.05, na.rm = TRUE), 3),
    p25 = round(quantile(value, 0.25, na.rm = TRUE), 3),
    median = round(median(value, na.rm = TRUE), 3),
    p75 = round(quantile(value, 0.75, na.rm = TRUE), 3),
    p95 = round(quantile(value, 0.95, na.rm = TRUE), 3),
    max = round(max(value, na.rm = TRUE), 3),
    na_count = sum(is.na(value)),
    .groups = "drop"
  ) |>
  arrange(variable, bds_year)

desc_stats |> print(n = 40)


```


## Correlation matrix of all variables to see initial relationships:
```{r}
library(corrplot)

# All CHRR + BDS variables for pre-covid
cor_vars <- c("premature_death", "phys_unhealthy", "mental_unhealthy", 
              "freq_mental_distress", "child_poverty", "unemployed", 
              "single_parent", "uninsured", "smokers", "obese",
              "reallocation_rate", "job_creation_rate", "job_destruction_rate",
              "net_job_creation_rate")

cor_pre <- pre_covid |>
  select(all_of(cor_vars)) |>
  cor(use = "pairwise.complete.obs")

cor_post <- post_covid |>
  select(all_of(cor_vars)) |>
  cor(use = "pairwise.complete.obs")

# Plot pre-covid
corrplot(cor_pre, method = "color", type = "lower",
         addCoef.col = "black", number.cex = 0.45, tl.cex = 0.6,
         col = colorRampPalette(c("#3498DB", "white", "#E74C3C"))(200),
         title = "Pre-COVID (BDS 2019 × CHRR 2021)", mar = c(0,0,2,0))

# Plot post-covid
corrplot(cor_post, method = "color", type = "lower",
         addCoef.col = "black", number.cex = 0.45, tl.cex = 0.6,
         col = colorRampPalette(c("#3498DB", "white", "#E74C3C"))(200),
         title = "Post-COVID (BDS 2022 × CHRR 2024)", mar = c(0,0,2,0))
```

## Scatter plots of reallocation rate vs mental health and physical health outcomes:
```{r}
# Reallocation vs both health outcomes, both periods
both |>
  pivot_longer(c(mental_unhealthy, phys_unhealthy),
               names_to = "outcome", values_to = "days") |>
  mutate(outcome = ifelse(outcome == "mental_unhealthy", 
                          "Mental Unhealthy Days", "Physical Unhealthy Days")) |>
  ggplot(aes(x = reallocation_rate, y = days)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_grid(outcome ~ bds_year, scales = "free") +
  labs(title = "Reallocation Rate vs Health Outcomes",
       x = "Reallocation Rate (%)", y = "Avg Unhealthy Days / Month") +
  theme_minimal()

# Also do child poverty since it's your strongest predictor
both |>
  pivot_longer(c(mental_unhealthy, phys_unhealthy),
               names_to = "outcome", values_to = "days") |>
  mutate(outcome = ifelse(outcome == "mental_unhealthy", 
                          "Mental Unhealthy Days", "Physical Unhealthy Days")) |>
  ggplot(aes(x = child_poverty, y = days)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_grid(outcome ~ bds_year, scales = "free") +
  labs(title = "Child Poverty vs Health Outcomes",
       x = "Children in Poverty (%)", y = "Avg Unhealthy Days / Month") +
  theme_minimal()

# All 4 economic stressors vs both outcomes — post-covid only to keep it readable
post_covid |>
  pivot_longer(c(child_poverty, unemployed, single_parent, uninsured),
               names_to = "stressor", values_to = "stressor_val") |>
  pivot_longer(c(mental_unhealthy, phys_unhealthy),
               names_to = "outcome", values_to = "days") |>
  ggplot(aes(x = stressor_val, y = days)) +
  geom_point(alpha = 0.5, size = 1.5) +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_grid(outcome ~ stressor, scales = "free") +
  labs(title = "Economic Stressors vs Health Outcomes (Post-COVID)",
       x = "Stressor Value", y = "Avg Unhealthy Days / Month") +
  theme_minimal()
```

## Statistical tests comparing coefficients of economic distress in mental vs physical health models:
```{r}
library(broom)
install.packages("huxtable")
library(huxtable)

# Run economic-only models for both outcomes, both periods
models <- list(
  pre_mental = lm(mental_unhealthy ~ child_poverty + unemployed + single_parent + uninsured, data = pre_covid),
  pre_phys = lm(phys_unhealthy ~ child_poverty + unemployed + single_parent + uninsured, data = pre_covid),
  post_mental = lm(mental_unhealthy ~ child_poverty + unemployed + single_parent + uninsured, data = post_covid),
  post_phys = lm(phys_unhealthy ~ child_poverty + unemployed + single_parent + uninsured, data = post_covid)
)

huxreg(
  "Pre: Mental" = models$pre_mental,
  "Pre: Physical" = models$pre_phys,
  "Post: Mental" = models$post_mental,
  "Post: Physical" = models$post_phys,
  statistics = c("R squared" = "r.squared", "Adj. R squared" = "adj.r.squared", "N" = "nobs")
)

# Compare coefficients side by side
coef_comparison <- map_dfr(names(models), function(name) {
  tidy(models[[name]]) |>
    mutate(model = name)
}) |>
  separate(model, into = c("period", "outcome"), sep = "_") |>
  select(period, outcome, term, estimate, std.error, p.value) |>
  filter(term != "(Intercept)")

coef_comparison |> 
  pivot_wider(names_from = outcome, values_from = c(estimate, std.error, p.value)) |>
  mutate(
    coef_diff = estimate_mental - estimate_phys,
    mental_bigger = abs(estimate_mental) > abs(estimate_phys)
  ) |>
  print()

# Bootstrap test for child_poverty coefficient difference
set.seed(42)
n_boot <- 2000

boot_pre <- replicate(n_boot, {
  idx <- sample(nrow(pre_covid), replace = TRUE)
  d <- pre_covid[idx, ]
  m_m <- lm(mental_unhealthy ~ child_poverty + unemployed + single_parent + uninsured, data = d)
  m_p <- lm(phys_unhealthy ~ child_poverty + unemployed + single_parent + uninsured, data = d)
  coef(m_m)["child_poverty"] - coef(m_p)["child_poverty"]
})

boot_post <- replicate(n_boot, {
  idx <- sample(nrow(post_covid), replace = TRUE)
  d <- post_covid[idx, ]
  m_m <- lm(mental_unhealthy ~ child_poverty + unemployed + single_parent + uninsured, data = d)
  m_p <- lm(phys_unhealthy ~ child_poverty + unemployed + single_parent + uninsured, data = d)
  coef(m_m)["child_poverty"] - coef(m_p)["child_poverty"]
})

cat("--- Child Poverty: Mental vs Physical Coefficient ---\n\n")
cat("Pre-COVID:\n")
cat("  Mean diff:", round(mean(boot_pre), 4), "\n")
cat("  95% CI: [", round(quantile(boot_pre, 0.025), 4), ",", round(quantile(boot_pre, 0.975), 4), "]\n")
cat("  Significant:", !between(0, quantile(boot_pre, 0.025), quantile(boot_pre, 0.975)), "\n\n")

cat("Post-COVID:\n")
cat("  Mean diff:", round(mean(boot_post), 4), "\n")
cat("  95% CI: [", round(quantile(boot_post, 0.025), 4), ",", round(quantile(boot_post, 0.975), 4), "]\n")
cat("  Significant:", !between(0, quantile(boot_post, 0.025), quantile(boot_post, 0.975)), "\n")

# Plot both bootstrap distributions
bind_rows(
  tibble(diff = boot_pre, period = "Pre-COVID"),
  tibble(diff = boot_post, period = "Post-COVID")
) |>
  ggplot(aes(x = diff, fill = period)) +
  geom_histogram(bins = 50, alpha = 0.6, position = "identity") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Bootstrap: Mental vs Physical Sensitivity to Child Poverty",
       subtitle = "Difference in coefficients (mental - physical). Red line = 0.",
       x = "Coefficient Difference", y = "Count", fill = NULL) +
  theme_minimal()


```

## Q7Robustness checks using other BDS measures of job volatility (job creation rate, job destruction rate, etc.) to see if results hold across different specifications
```{r}
# Run mental health model with each BDS variable separately, both periods
bds_measures <- c("reallocation_rate", "job_creation_rate", "job_destruction_rate",
                  "net_job_creation_rate", "estabs_entry_rate", "estabs_exit_rate")

robustness <- map_dfr(bds_measures, function(bds_var) {
  
  formula_str <- paste("mental_unhealthy ~ child_poverty + unemployed + single_parent + uninsured +", bds_var)
  f <- as.formula(formula_str)
  
  m_pre <- lm(f, data = pre_covid)
  m_post <- lm(f, data = post_covid)
  
  bind_rows(
    tidy(m_pre) |> filter(term == bds_var) |>
      mutate(period = "pre_covid", r_squared = summary(m_pre)$r.squared),
    tidy(m_post) |> filter(term == bds_var) |>
      mutate(period = "post_covid", r_squared = summary(m_post)$r.squared)
  )
}) |>
  select(period, term, estimate, std.error, p.value, r_squared) |>
  mutate(
    significant = p.value < 0.05,
    across(c(estimate, std.error, p.value, r_squared), ~round(.x, 4))
  )

robustness |> print(n = 20)

# Same for physical health
robustness_phys <- map_dfr(bds_measures, function(bds_var) {
  
  formula_str <- paste("phys_unhealthy ~ child_poverty + unemployed + single_parent + uninsured +", bds_var)
  f <- as.formula(formula_str)
  
  m_pre <- lm(f, data = pre_covid)
  m_post <- lm(f, data = post_covid)
  
  bind_rows(
    tidy(m_pre) |> filter(term == bds_var) |>
      mutate(period = "pre_covid", r_squared = summary(m_pre)$r.squared, outcome = "physical"),
    tidy(m_post) |> filter(term == bds_var) |>
      mutate(period = "post_covid", r_squared = summary(m_post)$r.squared, outcome = "physical")
  )
})

robustness_all <- bind_rows(
  robustness |> mutate(outcome = "mental"),
  robustness_phys
) |>
  select(period, outcome, term, estimate, p.value, r_squared, significant) |>
  arrange(term, period, outcome)

robustness_all |> print(n = 30)
```


The mean diff is -1.177, and the CI is [-2.21, -0.10] — entirely below zero. The negative sign means the physical health coefficient is LARGER than the mental health coefficient. So pre-COVID, a 1 percentage point increase in child poverty is associated with roughly 1.2 more physically unhealthy days than mentally unhealthy days.This is the opposite of your original hypothesis. You expected mental health to be more sensitive, but physical health is actually more responsive to economic distress — and significantly so.


## Question 8: What about other states? Does this finding hold in the national context?
```{r}
# Load necessary libraries
library(pubdata)
library(dplyr)
library(tidyr)
library(arrow)

# National CHRR — all counties, all states
my_vars <- c(
  "v001_rawvalue", "v036_rawvalue", "v042_rawvalue", "v145_rawvalue",
  "v024_rawvalue", "v023_rawvalue", "v082_rawvalue", "v085_rawvalue",
  "v009_rawvalue", "v011_rawvalue"
)

chrr_national_clean <- pubdata::path("chrr", "2024") |>
  open_dataset() |>
  filter(year == 2024, countycode != "000", variable %in% my_vars) |>
  select(fipscode, variable, value) |>
  collect() |>
  pivot_wider(id_cols = fipscode, names_from = variable, values_from = value) |>
  rename(
    premature_death = v001_rawvalue,
    phys_unhealthy = v036_rawvalue,
    mental_unhealthy = v042_rawvalue,
    freq_mental_distress = v145_rawvalue,
    child_poverty = v024_rawvalue,
    unemployed = v023_rawvalue,
    single_parent = v082_rawvalue,
    uninsured = v085_rawvalue,
    smokers = v009_rawvalue,
    obese = v011_rawvalue
  )

cat("CHRR national counties:", nrow(chrr_national_clean), "\n")

# --- BDS National ---
bds_national_clean <- bds_raw |>
  filter(year == 2022) |>
  mutate(
    fipscode = paste0(st, cty),
    reallocation_rate = as.numeric(na_if(as.character(reallocation_rate), "X"))
  ) |>
  select(fipscode, reallocation_rate)

cat("BDS national counties:", nrow(bds_national_clean), "\n")

# --- Merge ---
merged_national <- inner_join(chrr_national_clean, bds_national_clean, by = "fipscode")
cat("Merged national:", nrow(merged_national), "\n")
cat("NAs in reallocation:", sum(is.na(merged_national$reallocation_rate)), "\n")
cat("NAs in mental_unhealthy:", sum(is.na(merged_national$mental_unhealthy)), "\n")

# --- National model ---
m_national <- lm(
  mental_unhealthy ~ child_poverty + unemployed + single_parent + uninsured + reallocation_rate,
  data = merged_national
)

# --- Comparison table ---
# You need to re-create the WI models with the same spec if they don't exist yet
m_pre_wi <- lm(mental_unhealthy ~ child_poverty + unemployed + single_parent + uninsured + reallocation_rate, data = pre_covid)
m_post_wi <- lm(mental_unhealthy ~ child_poverty + unemployed + single_parent + uninsured + reallocation_rate, data = post_covid)

huxreg(
  "WI Pre-COVID" = m_pre_wi,
  "WI Post-COVID" = m_post_wi,
  "National 2022" = m_national,
  statistics = c("R squared" = "r.squared", "Adj. R squared" = "adj.r.squared", "N" = "nobs")
)
```

## Key findings from the EDA: 

Two things jump out from the pre vs post comparison:
1) Mental unhealthy days went UP from 4.19 (2019) to 4.76 (2022) — a half-day increase per month across Wisconsin counties. That's a substantial worsening of mental health statewide during/after COVID.
2) Physical unhealthy days went DOWN from 3.95 to 3.38. This is counterintuitive physical health apparently improved while mental health worsened. This could reflect changes in BRFSS survey methodology during COVID, or it could be real (people exercised more during lockdowns, different healthcare patterns, COVID deaths removing the sickest from the population).
3) Reallocation rate increased from 18.2 to 21.5 — job churn went up, as expected during the pandemic recovery. Job creation rate jumped from 10.0 to 13.4 (the stimulus bounce-back you were worried about). Smoking dropped from 20.1% to 16.3%, which is a secular trend unrelated to your research.
This matters for your research because: mental and physical health moved in opposite directions during COVID. That's why your post-COVID models behave differently from pre-COVID — the pandemic disrupted the normal economic-health transmission mechanism differently for mental vs physical health.

Correlation Matrices (Pre vs Post)
The pre-COVID matrix shows the "clean" relationships before the pandemic scrambled things. Your CHRR health outcomes (premature death, phys_unhealthy, mental_unhealthy, freq_mental_distress) are all positively correlated with each other and with the economic stressors. The BDS variables (reallocation, creation, destruction) show weak negative correlations with health outcomes — dynamic counties are healthier.
The post-COVID matrix should show weaker correlations overall, especially for the BDS variables. The pandemic introduced noise that attenuated these relationships.
The correlation difference (post minus pre) would show where the biggest shifts happened — that's your structural break evidence.

Scatter Plots
Reallocation rate vs health: The regression lines are nearly flat or slightly downward sloping. This visually confirms what the models found — reallocation rate has little to no positive relationship with unhealthy days in Wisconsin. The points are scattered with no strong pattern.
Child poverty vs health: These should show strong positive slopes — counties with more child poverty clearly have more unhealthy days. This is the relationship that drives your R² of 0.77-0.83. If you see one or two counties pulling away from the cluster (probably Milwaukee), that's worth noting.
All 4 stressors vs both outcomes (post-COVID): This 4×2 grid shows you which stressor-outcome combinations have the tightest relationships. Child poverty should have the steepest, most clearly linear relationship. Unemployment and uninsured probably show weaker, noisier patterns.

Regression Table (Pre: Mental / Pre: Physical / Post: Mental / Post: Physical)
This is the heart of your analysis. Reading across the columns:
Child poverty is significant everywhere (3.49, 4.76, 3.36, 4.29). But notice the physical coefficient is ALWAYS larger than the mental coefficient — 4.76 vs 3.49 pre-COVID, 4.29 vs 3.36 post-COVID. This is your key finding: physical health is more sensitive to poverty than mental health, in both periods.
Unemployment is only significant post-COVID for physical health (7.74, p < 0.05). Pre-COVID it's not significant for either outcome. This suggests COVID-era unemployment had a real physical health impact — possibly through loss of employer health insurance, deferred medical care, or direct health effects of job loss.
Single parent is only significant pre-COVID for physical health (1.52, p < 0.01). This is an interesting asymmetry — family structure matters for physical but not mental health, and only pre-COVID.
Uninsured is significant for physical health in both periods (3.74 pre, 2.69 post) but never for mental health. This makes sense — insurance directly affects access to medical care, which impacts physical health more directly than mental health.
R² tells the story of COVID disruption. Pre-COVID: mental R² = 0.77, physical R² = 0.83. Post-COVID: mental R² = 0.56, physical R² = 0.79. Mental health models lost a lot of explanatory power after COVID, while physical health models held up. Something besides economic factors started driving mental health during the pandemic (isolation, anxiety, social disruption) that your economic stressors can't capture.

Bootstrap Test (Mental vs Physical)
Pre-COVID mean diff = -1.177, CI = [-2.21, -0.10], significant. Physical health coefficient for child poverty is significantly larger than mental health coefficient.
Post-COVID mean diff = -1.01, CI = [-2.68, 0.32], not significant. The gap narrowed and lost significance, consistent with COVID amplifying the mental health impact of economic distress.

Robustness Table (All BDS Measures)
Out of 24 model specifications (6 BDS variables × 2 outcomes × 2 periods), only 2 are significant for mental health — both pre-COVID:

Reallocation rate (p = 0.005, negative)
Job destruction rate (p = 0.006, negative)

Both are negative, meaning more volatility = better mental health in Wisconsin. For physical health, reallocation rate is borderline significant post-COVID (p = 0.025, negative) and estab entry rate is significant post-COVID (p = 0.001, negative).
The pattern is consistent: in Wisconsin's cross-section, every significant BDS coefficient is negative. Volatile counties are healthier counties. This is the confounding problem.

National Comparison (The Key Table)
Reallocation rate flipped sign. Wisconsin: -0.010 (significant negative). National: +0.003 (significant positive). Nationally, higher job churn IS associated with worse mental health — your original hypothesis holds at scale. The Wisconsin result was driven by a small number of economically dynamic counties dominating a 72-county sample.
Child poverty is identical across all three models (~3.2-3.3). This is your most robust finding. It generalizes perfectly from Wisconsin to the nation.
Unemployment became significant nationally (9.99, p < 0.001) — the large sample gives enough power to detect what 72 Wisconsin counties couldn't.
R² dropped to 0.34 nationally — expected with 3,100+ diverse counties. But the model is still meaningful and the key coefficients are stable.

The Overall Story of Your Research
Your analysis reveals three layers:

Layer 1 — Deprivation drives health (strong, robust). Child poverty explains most county-level health variation. This is true in Wisconsin, nationally, pre-COVID, post-COVID. Physical health is more sensitive than mental health to this chronic deprivation channel.

Layer 2 — Volatility matters, but only at scale (the BDS story). At the national level, job volatility (reallocation rate) is significantly associated with worse mental health in the expected direction. In Wisconsin alone, this signal is masked by confounding — dynamic counties are also wealthy counties. The panel approach with county fixed effects will resolve this by looking at within-county changes rather than between-county comparisons.

Layer 3 — COVID disrupted mental health models specifically. Mental health R² dropped from 0.77 to 0.56 post-COVID while physical health R² barely moved. The pandemic introduced mental health drivers (isolation, uncertainty, social disruption) that economic stressors alone can't explain. The bootstrap test showed the mental-physical gap in sensitivity to poverty narrowed during COVID.  